---
title: "MA615 Midterm Project"
author: "Ziyan Li"
date: "October 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(vcd)
```

```{r}
# read data
hms <- read.csv("HMS raw data 2016.csv")
```

# 1. Dataset Overview
This dataset is from MSSP consulting project, "Healthy Minds Study", which focuses on BU student mental health and related issues. The dataset has 532 variables and 2551 observations, most of which are NAs. Since it is a survey data, there are different types of data. It is necessary to take a overview at the dataset.
```{r}
# write a function to help us recognize data type for each column
describe <- function(x){
  y = 0
  if (sum(is.na(x)) == length(x)) {
    y = "only NA"
  }
  else {
    x = na.omit(x)
    if (is.numeric(unique(x[2])) == FALSE){
      y = "text"
    }
    else{
      if (min(x) == 0 & max(x) == 1){
        y = "0 & 1"
      }
      else if (min(x) == 1 & max(x) == 1){
        y = "1 & NA"
      }
      else{
        if (max(x) < 10){
          y = "scale/categorical"
        }
        else{
          y = "numeric"
        }
      }
    }
  }
  return(y)
}
```
Apply the function to our dataset and return a statistics table.
```{r, results='asis'}
overview <- as.data.frame(sapply(hms, describe))
names(overview) <- "type"
kable(overview %>% group_by(type) %>% summarise(count=length(type)))
```
All these types are mixed within those 532 columns. So we have to split the whole dataset and reorganize them based the codebook we have in order to make our future analysis convenient.

# 2. Reorganize Data

```{r}
row.names(hms) <- hms$responseid
# absorb key words of column names
key <- paste("df",unlist(lapply(strsplit(names(hms), split="_"),function(x){x[1]})),sep = "_")
```

Create a function to help us split our dataset into separate data frames.
```{r}
pull <- function(x){   # input x is a single key word
  df = as.data.frame(hms[,grep(x,key)]) # search in column names
  colnames(df) = names(hms)[grep(x,key)] 
  rownames(df) = rownames(hms)  # if not defined, single column section will lose row names
  assign(x, df, envir=.GlobalEnv) # save data frame to our global environment
}
```

Apply the function to our dataset.
```{r}
invisible(lapply(unique(key),pull)) 
# around 100 separate data frames will be created in global environment
```

Some adjustment will be needed.
```{r}
df_sex <- data.frame(df_gender, df_sex, df_relship)
df_socioeco <- data.frame(df_fincur, df_finpast, df_educ)
df_academic <- data.frame(df_degree, df_transfer, df_yr, df_enroll, df_field, df_gpa, df_aca, df_timeclass, df_timestud, df_persist)
df_depress <- data.frame(df_phq9, df_deprawsc, df_dep$dep_impa, df_phq2)
names(df_depress)[11] <- "dep_impa"
df_anxiety <- data.frame(df_gad7, df_anx$anx_score)
names(df_anxiety)[9] <- "anx_score"
df_eating <- data.frame(df_thin, df_body, df_height, df_weight, df_scoff, df_ed$ed_scoff)
names(df_eating)[11] <- "ed_scoff"
df_violence <- data.frame(df_violence, df_assault)
df_use <- data.frame(df_alc, df_binge, df_smok, df_drug)
df_help <- data.frame(df_med$med_help, df_ther$ther_help)
names(df_help) <- c("med_help", "ther_help")
df_stigma <- data.frame(df_stig$stig_pcv_1, df_stig$stig_per_1)
names(df_stigma) <- c("stig_pcv_1", "stig_per_1")
df_use_ct <- data.frame(df_ther$ther_any, df_ther$ther_vis, df_ther$ther_cur, df_prov)
names(df_use_ct)[1:3] <- c("ther_any", "ther_vis", "ther_cur")
df_satisf <- data.frame(df_sat, df_ther$ther_helped_me)
names(df_satisf)[length(df_satisf)] <- "ther_helped_me"
df_medic <- data.frame(df_meds, df_why)
df_informal <- data.frame(df_inf, df_talk2, df_talkaca, df_talksup)
df_know_mit <- data.frame(df_know$know_sp, df_dep[,3:10], df_anx[,2:5], df_ea, df_gk)
names(df_know_mit)[1] <- "know_sp"
df_know_cs <- data.frame(df_know$knowwher, df_outreach, df_heard, df_camp)
names(df_know_cs)[1] <- "knowwher"
df_belief <- data.frame(df_meds$meds_help_me, df_ther$ther_help_me)
names(df_belief) <- c("meds_help_me", "ther_help_me")
df_identity <- data.frame(df_self, df_dep$dep_secret, df_men)
names(df_identity)[2] <- "dep_secret"
df_stig_pcv <- data.frame(df_stig$stig_pcv_2, df_stig$stig_pcv_3, df_stig$stig_pcv_1)
names(df_stig_pcv) <- c("stig_pcv_2", "stig_pcv_3", "stig_pcv_1")
df_stig_per <- data.frame(df_stig$stig_per_2, df_stig$stig_per_3, df_stig$stig_per_1)
names(df_stig_per) <- c("stig_per_2", "stig_per_3", "stig_per_1")
df_belong <- data.frame(df_belong1, df_belong2, df_belong3, df_belong4, df_belong5, df_belong6, df_belong7, df_belong8, df_belong9)
df_camp_cli <- data.frame(df_mh, df_cli, df_st$st_promote, df_admin, df_env)
names(df_camp_cli)[4] <- "st_promote"
df_diverse <- data.frame(df_divers, df_discrim)
```

Finally, save our global environment as .RDa file.
```{r}
save.image("split_data.RData")
head(ls(), 20) # around 100 separate data.frames in total
```
The output of reorganization is shown in the pdf file.

# 3. Clean Text Input
First of all, let's take a overview of all the text columns.
```{r}
text <- as.data.frame(hms[,grep("text",colnames(hms))])
text[text==""] <- NA
colSums(!is.na(text))
```
There are 36 text columns 4 of which are empty. According to our client's demand, we firstly focus on clean the text input related to demographic variables. Here is an example our text cleaning workflow using 4 demographic variables.

#### (1) clean race text input
```{r}
# clean race
df_race$race_other_text[df_race$race_other_text==""] <- NA
colSums(!is.na(df_race)) # all colums related to race question.
race_text <- unique(hms$race_other_text)
mixed <- c("Bi","bi","ix","and","-","part")
text_mixed <- race_text[grep(paste(mixed, collapse = "|"), race_text)]
asian <- c("sian","ndia","akist","nese")
text_asian <- race_text[grep(paste(asian, collapse = "|"), race_text)]
df_race$race[df_race$race_white == 1] <- "white"
df_race$race[df_race$race_black == 1] <- "black"
df_race$race[df_race$race_asian == 1 | df_race$race_other_text %in% text_asian] <- "asian"
df_race$race[rowSums(df_race[2:8], na.rm = TRUE) >1 | df_race$race_other_text %in% text_mixed] <- "mixed"
df_race$race[rowSums(!is.na(df_race[2:10])) >0 & is.na(df_race$race)] <- "other"
# out put: a single variable containing race information
kable(df_race %>% group_by(race) %>% summarise(count=length(race)))
```

### (2) clean degree text input
```{r}
# clean degree
df_degree$degree_other_text[df_degree$degree_other_text==""] <- NA
colSums(!is.na(df_degree))
unique(df_degree$degree_other_text[rowSums(is.na(df_degree[,2:6]))==5])
df_degree$degree[rowSums(!is.na(df_degree)) >0] <- "graduate"
df_degree$degree[df_degree$degree_bach==1] <- "undergraduate"
df_degree$degree[rowSums(df_degree[,2:5], na.rm = TRUE) >0] <- "graduate" # record the highest possible degree
kable(df_degree %>% group_by(degree) %>% summarise(count=length(degree)))
```

### (3) clean gender text input
```{r}
# clean gender
# for gender we can simply ignore text input since our client asked to recode gender into four categories.
df_gender$gender1[df_gender$gender == 1] <- "male"
df_gender$gender1[df_gender$gender == 2] <- "female"
df_gender$gender1[df_gender$gender == 3 | df_gender$gender == 4] <- "trans"
df_gender$gender1[df_gender$gender == 5 | df_gender$gender == 6] <- "other"
kable(df_gender %>% group_by(gender1) %>% summarise(count=length(gender1)))
```

### (4) clean citizen text input
```{r}
# clean citizen
df_citizen$citizen1[df_citizen$citizen == 1] <- "domestic"
df_citizen$citizen1[df_citizen$citizen == 0] <- "international"
kable(df_citizen %>% group_by(citizen1) %>% summarise(count=length(citizen1)))
```

#### combine these four new columns into a new data.frame
```{r}
# combine demographic
df_demo <- data.frame(gender=df_gender$gender1, race=df_race$race, citizen=df_citizen$citizen1, degree=df_degree$degree)
```

# 4. Plot Data
Since most of the efforts are put into cleaning process, the visualization part will be a little plain.
```{r}
par(mfrow=c(2,2))
demo1 <- table(df_demo$race, df_demo$degree)
demo1
demo2 <- table(df_demo$gender, df_demo$degree)
demo2
demo3 <- table(df_demo$degree, df_demo$citizen)
demo3
demo4 <- table(df_demo$race, df_demo$citizen)
demo4
mosaicplot(demo1, color = c("aquamarine4","antiquewhite3"), main="race & degree", xlab="race", ylab="degree")
mosaicplot(demo2, color = c("salmon","tomato4"), main="gender & degree", xlab="gender", ylab="degree")
mosaicplot(demo3, color = c("steelblue4","steelblue2"), main="degree & citizen", xlab="degree", ylab="citizen")
mosaicplot(demo4, color = c("bisque3","lightpink4"), main="race & citizen", xlab="race", ylab="citizen")
```

